---
title: "Average-Forecasting-Metrics"
author: "Amanda Bleichrodt"
date: "2024-01-29"
output: html_document
---
```{r needed-packages}
library(tidyverse)
```

To compare forecasting metrics across locations, models and forecasting horizons
this file calculates the average metrics across all forecasting periods. The
data is grouped by location, model, calibration period and forecasting horizon.

However, first the data is read in from the working directory. The name of the 
file should be "all-metrics.csv", the file created from the "Merging-Forecast-
Period-Metrics.Rmd" code. 

# Reading in multi-period forecasting metrics 
```{r reading-in-metrics}
###################################################
# Reading in the forecast-period specific metrics #
###################################################
all_metrics <- read_csv("~/Georgia State R/MonkeyPox/Mpox Metrics Paper/Forecast-Metrics/all-metrics.csv")
```

Once the data is read into R, the code then calculates the overall average
metrics across forecasting dates for each location, model, calibration period,
and forecasting horizon. The average metrics are then exported to the working
directory as a '.csv'. 

# Calculating the overall average metrics 
```{r average-metrics}
###############################
# Calculating average metrics #
###############################
averageMetrics <- all_metrics %>%
  dplyr::group_by(Location, Model, Calibration, Horizon) %>% # Grouping variables
  dplyr::mutate(AvgMSE = mean(MSE), # Average MSE
                AvgMAE = mean(MAE), # Average MAE
                AvgPI = mean(PI), # Average 95% PI
                AvgWIS = mean(WIS)) %>% # Average WIS
  dplyr::distinct(Location, Model, Calibration, Horizon, .keep_all = T) %>% # Keeping unique rows
  dplyr::select(Location, Model, Calibration, Horizon, AvgMSE, AvgMAE, AvgPI, AvgWIS) # Needed variables

##############################
# Saving the average metrics #
##############################
write.csv(averageMetrics, file = "~/Georgia State R/MonkeyPox/Mpox Metrics Paper/Forecast-Metrics/average-metrics-overall.csv")
```

As we are looking at multiple countries, another view of interest is looking 
at the average metric by country specific epidemic phase. Here, we look at 
for phases of interest: (1) Ascending, (2) Peak, (3) Descending, (4) tail-end.
Once the average metrics are calculated, the file is then exported as a '.csv'
to the working directory.

# Calculating the time-period specific average metrics
```{r time-period-metrics}
###############################################
# Average period specific performance metrics #
###############################################
avg_metric_date <- all_metrics %>%
  dplyr::mutate(Date = anytime::anydate(Date), # Changing the date format to "Date"
               
                ############################
                # Ascending phases - Dates #
                ############################
                Aphase = case_when(
                  Location == "Germany" ~ '2022-06-23', # Germany
                  Location ==  "Spain" ~ '2022-06-30', # Spain 
                  Location == "Canada" ~ '2022-06-30', # Canada
                  Location == "United Kingdom" ~ '2022-07-07', # UK
                  Location == "France" ~ '2022-07-14', # France
                  Location == "US(CDC)" ~ '2022-07-21', # US-CDC
                  Location == "Brazil" ~ '2022-07-28', # Brazil
                  Location == "World" ~ '2022-07-28', # World
                  Location == "US(OWID)" ~ '2022-08-04' # US-OWID
                ),
                
                ############################
                # Peak phases - Start date #
                ############################
                PSphase = case_when(
                  Location == "Germany" ~ '2022-06-30', # Germany
                  Location == "France" ~ '2022-07-21', # France
                  Location == "Spain" ~ '2022-07-07', # Spain
                  Location == "Canada" ~ '2022-07-07', # Canada
                  Location == "Brazil" ~ '2022-08-04', # Brazil
                  Location == "World" ~ '2022-08-04', # World
                  Location == "US(CDC)" ~ '2022-07-28', # US-CDC
                  Location == "US(OWID)" ~ '2022-08-11', # US-OWID
                  Location == "United Kingdom" ~ '2022-07-14' # UK
                ),
                
                ##########################
                # Peak phases - End date #
                ##########################
                PEphase = case_when(
                  Location == "Germany" ~ '2022-07-14', # Germany
                  Location == "France" ~ '2022-08-25', # France
                  Location == "Spain" ~ '2022-08-11', # Spain
                  Location == "United Kingdom" ~ '2022-7-14', # UK
                  Location == "US(OWID)" ~ '2022-08-11', # US-OWID
                  Location == "US(CDC)" ~ '2022-07-28', # US-CDC
                  Location == "World" ~ '2022-08-11', # World
                  Location == "Canada" ~ '2022-07-28', # Canada
                  Location == "Brazil" ~ '2022-09-08' # Brazil
                ), 
                
                ###########################
                # Descending - Start date #
                ###########################
                DSphase = case_when(
                  Location == "Germany" ~ '2022-07-21', # Germany
                    Location == "France" ~ '2022-09-01', # France
                    Location == "Spain" ~ '2022-08-18', # Spain
                    Location == "United Kingdom" ~ '2022-07-21', # UK
                    Location == "US(OWID)" ~ '2022-08-18', # US-OWID
                    Location == "US(CDC)" ~ '2022-08-04', # US-CDC
                    Location == "World" ~ '2022-08-18', # World
                    Location == "Canada" ~ '2022-08-04', # Canada
                    Location == "Brazil" ~ '2022-09-15' # Brazil
                ),
                
                #########################
                # Descending - End date #
                #########################
                DEphase = case_when(
                  Location == "Germany" ~ '2022-10-13', # Germany
                    Location == "France" ~ '2022-10-13', # France
                    Location == "Spain" ~ '2022-10-06', # Spain
                    Location == "United Kingdom" ~ '2022-10-13', # UK
                    Location == "US(OWID)" ~ '2022-12-01', # US-OWID
                    Location == "US(CDC)" ~ '2022-11-17', # US-CDC
                    Location == "World" ~ '2022-12-08', # World
                    Location == "Canada" ~ '2022-10-13', # Canada
                    Location == "Brazil" ~ '2022-11-17' # Brazil
                ),
                
                #####################################
                # Creating the time indicator (1-4) #
                #####################################
                timeIndicator = ifelse(Date <= anytime::anydate(Aphase), 1, 
                                       ifelse(Date <= anytime::anydate(PEphase) & Date >= anytime::anydate(PSphase), 2,
                                              ifelse(Date <= anytime::anydate(DEphase) & Date >= anytime::anydate(DSphase), 3, 
                                                     ifelse(Date > anytime::anydate(DEphase), 4, NA))))) %>%
  ######################
  # Group-by variables #
  ######################
  dplyr::group_by(Location, Model, Horizon, Calibration, timeIndicator) %>%
  dplyr::mutate(avgMSE = mean(MSE), # Average MSE
                avgMAE = mean(MAE), # Average MAE
                avg95PI = mean(`PI`), # Average 95% PI
                avgWIS = mean(WIS),  # Average WIS
                timeIndicator = as.character(timeIndicator)) %>% # Saving time indicator as character
  dplyr::distinct(Location, Model, Horizon, Calibration, timeIndicator, .keep_all = T) %>% # Keeping unique rows 
  dplyr::select(Model, Location, Horizon, Calibration, timeIndicator, avgMSE, avgMAE, avg95PI, avgWIS) # Needed variables


##########################################
# Saving the average time-period metrics #
##########################################
write.csv(avg_metric_date, file = "~/Georgia State R/MonkeyPox/Mpox Metrics Paper/Forecast-Metrics/phase-specific-metrics.csv")

```

