---
title: "Combining-Performance-Metrics"
author: "Amanda Bleichrodt"
date: "2024-01-28"
output: html_document
---

## Needed packages
```{r needed-packages}
library(qdapRegex)
library(tidyverse)
library(stringr)
```

## About

This file reads in the forecast performance metrics for each calibration 
period, forecast horizon, location, and model and combines them into one single
'.csv' file. The '.csv' file is then exported to the working directory.

```{r empty-vector}
#############################
# Creating the empty vector #
#############################
finalList <- NA
```

## Reading in the performance metrics

Prior to combining the forecast metrics, we need to read them into R. This
section creates a vector of file names. The file names are pulled from the 
'Forecast-Metrics' folder. However, the names of the folders themselves will
also be read into R; this will be handled in the next step of the code.

```{r reading-in-metrics}
allMetrics <- list.files (path = "~/Georgia State R/MonkeyPox/Mpox Metrics Paper/Forecast-Metrics/", # Folder with saved forecasts
                          recursive = T, # Calling all files
                          full.names = T) # Telling R we want the full file names 
```

## Combining the metrics

Now that the file names are read into R, we will loop through each metric file,
formatted as needed, and combine them into one file that will be exported.

```{r combining-loop}
for(i in 1:length(allMetrics)){
  
  # File name
  metricName <- allMetrics[i]
  
  ################################################
  # Pulling necessary information from file name #
  ################################################
  
  # Model framework
  model_framework <- qdapRegex::ex_between(metricName, "Metrics/", "/")[[1]][1]
  
  # Model type
  if(model_framework %in% c("ARIMA", "GAM", "SLR", "Prophet")){
    
    model_type <- qdapRegex::ex_between(metricName, paste0(model_framework, "/"), "-weekly")[[1]][1]
    
  }else{
    
    model_type <- qdapRegex::ex_between(metricName, "performance-", "-onset")[[1]][1]
    
  }
  
  # Calibration period
  calibration <- qdapRegex::ex_between(metricName, "calibration-", "-")[[1]][1]
  
  # Forecast horizon
  horizon <- qdapRegex::ex_between(metricName, "horizon-", "-")[[1]][1]
  
  # Forecast date 
  forecastDate <- str_sub(metricName,-14,-5)

  
  ###################
  # Creating labels #
  ###################
  
  # Model type 
  modelLabel <- switch (model_framework,
    "n-sub-epidemic" = "SE-",
    "Spatial-wave" = "SW-",
    model_type
  )
  
  # Weight label
  if(grepl("Unweighted", model_type)){
    
    weightLabel <- "-UW"
    
    modelFinal <- "Ensemble"
    
  }else if(grepl("Weighted", model_type)){
    
    weightLabel <- "-W"
    
    modelFinal <- "Ensemble"
    
  }

  ################################
  # Reformatting the metric file #
  ################################
  
  # Reading in the file
  data <- read.csv(metricName, header = T)
  
  ########################################################################
  # Fixing the model name and final data file - ARIMA, GLM, GAM, Prophet #
  ########################################################################
  if(model_framework %in% c("ARIMA", "GAM", "SLR", "Prophet")){
    
    # Location 
    location <- qdapRegex::ex_between(metricName, paste0("horizon-", horizon, "-"), paste0("-", forecastDate))[[1]][1]
  
    # Creating the data to be merged 
    dataFinal <- data %>%
      dplyr::mutate(Model = model_type, # Model type 
                     Location  = location, # Location
                     Calibration = calibration, # Calibration period 
                     Horizon = horizon, # Forecasting horizon 
                     Date = forecastDate) %>% # Forecasting Date
       dplyr::rename("PI" = "X95PI") %>% # Renaming the 95% PI
       dplyr::select(Location, Model, Calibration, Horizon, Date, MSE, MAE, PI, WIS) # Needed variables 
       
  ########################################################
  # Fixing the model name and final data file - Ensemble #
  ########################################################
  }else if(grepl("Ensemble", model_type)){
    
    # Location 
    location <- qdapRegex::ex_between(metricName, paste0("calibration-", calibration, "-"), paste0("-", forecastDate))[[1]][1]
  
    # Creating the data to be merged 
    dataFinal <- data %>%
     dplyr::mutate(Model = paste0(modelLabel, modelFinal, "(", row_number() + 1, ")", weightLabel), # Model type
                   Location = location, # Location
                   Calibration = calibration, # Calibration period 
                   Horizon = horizon, # Forecasting horizon 
                   Date = forecastDate) %>% # Forecast date
      dplyr::rename("PI" = "Coverage.95.PI") %>% # 95% PI
      dplyr::select(Location, Model, Calibration, Horizon, Date, MSE, MAE, PI, WIS) # Needed variables
    
  ######################################################
  # Fixing the model name and final data file - Ranked #
  ######################################################
  }else{
    
    # Location 
    location <- qdapRegex::ex_between(metricName, paste0("calibration-", calibration, "-"), paste0("-", forecastDate))[[1]][1]
  
    # Creating the data file to be merged 
    dataFinal <- data %>%
     dplyr::mutate(Model = paste0(modelLabel, "Ranked(", row_number(), ")"), # Model type 
                  Location = location, # Location 
                   Calibration = calibration, # Calibration period 
                   Horizon = horizon, # Horizon 
                   Date = forecastDate) %>% # Location
      dplyr::rename("PI" = "Coverage.95.PI") %>% # 95% PI
      dplyr::select(Location, Model, Calibration, Horizon, Date, MSE, MAE, PI, WIS) # Needed variables 
    
    }
  
  #####################################
  # Adding the data to the final list #
  #####################################
  finalList <- rbind(finalList, dataFinal)
  
}
```

# Fixing country names 
```{r fixing-names}
############################
# Adjusting Location Names #
############################
finalList <- finalList %>%
  dplyr::mutate(Location = ifelse(Location == "US-OWID", "US(OWID)",
                                  ifelse(Location == "US-CDC", "US(CDC)",
                                         ifelse(Location == "United-Kingdom", "United Kingdom", Location))))
```

## Saving the final metrics

Now that the metrics have been combined, the top row of NAs needs to be removed.
The NAs occur because the initial state of the 'finalList' is NA. Once removed,
the file will be saved as a '.csv' to the working directory.

```{r saving-file}
###############################
# Removing the top row of NAs #
###############################
listToExport <- na.omit(finalList)

#####################
# Exporing the list #
#####################
write.csv(listToExport, "~/Georgia State R/MonkeyPox/Mpox Metrics Paper/Forecast-Metrics/all-metrics.csv", row.names = F)
```

