---
title: "Calculating-Skill-Scores"
author: "Amanda Bleichrodt"
date: "2024-01-29"
output: html_document
---

# Needed packages 
```{r needed-packages}
library(tidyverse)
```

# Reading in source data
```{r Reading-In-Metrics}
#######################################
# Reading in the average metrics data #
#######################################
average_metrics <- read_csv("~/Georgia State R/MonkeyPox/Mpox Metrics Paper/Forecast-Metrics/average-metrics-overall.csv") %>%
  # Fixing location labels 
  dplyr::mutate(Location = ifelse(Location == "US-CDC", "US(CDC)",
                                  ifelse(Location == "US-OWID", "US(OWID)",
                                         ifelse(Location == "United-Kingdom", "United Kingdom", Location))))

#############################
# Winkler scores for 95% PI #
#############################
average_metricsPI <- read_csv("~/Georgia State R/MonkeyPox/Mpox Metrics Paper/Forecast-Metrics/Average-winkler-Scores.csv") %>%
  # Fixing location labels 
  dplyr::mutate(Location = ifelse(Location == "US-CDC", "US(CDC)",
                                  ifelse(Location == "US-OWID", "US(OWID)",
                                         ifelse(Location == "United-Kingdom", "United Kingdom", Location))))
```

# Preparing for calculating the skill scores
```{r preparing-data}
#############################################################
# Creating an empty data frame to fill in with skill-scores #
#############################################################
fullMergeData <- NA
```

# Calculating the skill scores

Skill scores provide a way of quantifying the improvment of one model over 
another. In this case, we are using the established ARIMA model as our baseline
and comparing it to the *n*-sub-epidemic and spatial-wave framework models. The
equation for skill scores is as follow:

         Baseline Model - Comparison Model / Baseline * 100.
```{r looping-through-horizons}
#############################
# Starting the horizon loop #
#############################
for(h in 1:4){
  
  ######################################
  # Sub-setting only the ARIMA metrics #
  ######################################
  ARIMAdata3 <- average_metrics[, -1] %>%
    dplyr::filter(Calibration == 11, # Filtering to include only 11-week calibration 
                  Model == "ARIMA", # Filtering to include only ARIMA metrics 
                  Horizon == h) %>% # Filtering to include only indexed forecasting horizon 
    dplyr::rename(ARIMA.MSE = AvgMSE, # Renaming the MSE metric for ARIMA 
                  ARIMA.MAE = AvgMAE, # Renaming the MAE metric for ARIMA 
                  ARIMA.WIS = AvgWIS) %>% # Renaming the WIS metric for ARIMA 
    dplyr::select(-Calibration, -Model, -AvgPI) # Selecting the needed variables 
  
  # Handling the ARIMA Winkler scores
  ARIMADataWinkler <- average_metricsPI %>%
    dplyr::filter(Model == "ARIMA", # Filtering to include only ARIMA metrics
                  Horizon == h) %>% # Filtering to include only indexed forecasting horizon 
    dplyr::rename(ARIMA.Winkler = meanWS) %>% # Renaming the Winkler score for ARIMA
    dplyr::select(Location, Horizon, ARIMA.Winkler) # Selecting the needed variables
  
  # Combining the two data frames
  ARIMAdata <- merge(ARIMAdata3, ARIMADataWinkler)
  
  ################################################################
  # Sub-setting only the n-sub-epidemic and spatial-wave metrics #
  ################################################################
  modelsOfInterest3 <- average_metrics[, -1] %>%
    dplyr::filter(Calibration == 11, # Filtering to include only 11-week calibration 
                  Model != "GAM", # Removing the GAM metrics 
                  Model != "SLR", # Removing the SLR metrics 
                  Model != "Prophet", # Removing the Prophet metrics 
                  Model != "ARIMA", # Removing the ARIMA metrics 
                  Horizon == h) %>%  # Filtering to include only indexed forecasting horizon 
    dplyr::select(-Calibration, -AvgPI) # Selecting the needed variables 
  
  # Handling the Other Model Winkler scores
  OtherDataWinkler <- average_metricsPI %>%
      dplyr::filter(Model != "GAM", # Removing the GAM metrics 
                    Model != "SLR", # Removing the SLR metrics 
                    Model != "Prophet", # Removing the Prophet metrics 
                    Model != "ARIMA", # Removing the ARIMA metrics 
                    Horizon == h) %>%  # Filtering to include only indexed forecasting horizon 
    dplyr::rename(AvgWS = meanWS) %>% # Renaming the Winkler score for ARIMA
    dplyr::select(Location, Horizon, Model, AvgWS) # Selecting the needed variables
  
  # Combining the two data frames
  modelsOfInterest <- merge(modelsOfInterest3, OtherDataWinkler)
  
  ################################
  # Calculating the skill-scores #
  ################################
  calculatingSkillScores <- right_join(ARIMAdata, modelsOfInterest) %>% # Combining the ARIMA and other metrics 
    dplyr::mutate(MSE.Score = ((ARIMA.MSE-AvgMSE)/ARIMA.MSE)*100, # MSE skill scores 
                  MAE.Score = ((ARIMA.MAE-AvgMAE)/ARIMA.MAE)*100, # MAE skill scores 
                  WIS.Score = ((ARIMA.WIS-AvgWIS)/ARIMA.WIS)*100, # WIS skill scores 
                  PI.Score = ((ARIMA.Winkler-AvgWS)/ARIMA.Winkler)*100) %>% # PI skill scores 
    dplyr::select(Location, Model, Horizon, MSE.Score, MAE.Score, PI.Score, WIS.Score) # Selecting the needed variables 
  
  ####################################
  # Merging each forecasting horizon #
  ####################################
  fullMergeData <- rbind(fullMergeData, calculatingSkillScores) %>% # Merging each forecasting horizon 
    stats::na.omit() %>% # Removing the NAs
    dplyr::arrange(Horizon) # Ordering data from smallest to largest forecasting horizon 
  
}# End of loop going through forecasting horizons 
```
# Preparing the data tables 
```{r preparing-data-tables}
###############
# MSE Metrics #
###############
MSE.Data <- fullMergeData %>%
  dplyr::select(Location, Model, Horizon, MSE.Score) %>% # Selecting MSE variables 
  dplyr::mutate(Metric = "MSE") %>% # Adding metric indicator 
  tidyr::pivot_wider(names_from = Model, values_from = `MSE.Score`) # Long-to-wide data

###############
# MAE Metrics #
###############
MAE.Data <- fullMergeData %>%
  dplyr::select(Location, Model, Horizon, MAE.Score) %>% # Selecting MAE variables 
  dplyr::mutate(Metric = "MAE") %>% # Adding metric indicator 
  tidyr::pivot_wider(names_from = Model, values_from = `MAE.Score`) # Long-to-wide data

##############
# PI Metrics #
##############
PI.Data <- fullMergeData %>%
  dplyr::select(Location, Model, Horizon, PI.Score) %>% # Selecting MAE variables 
  dplyr::mutate(Metric = "PI") %>% # Adding metric indicator 
  tidyr::pivot_wider(names_from = Model, values_from = `PI.Score`) # Long-to-wide data

###############
# WIS Metrics #
###############
WIS.Data <- fullMergeData %>%
  dplyr::select(Location, Model, Horizon, WIS.Score) %>% # Selecting WIS variables 
  dplyr::mutate(Metric = "WIS") %>% # Adding metric indicator 
  tidyr::pivot_wider(names_from = Model, values_from = `WIS.Score`) # Long-to-wide data

#############################################
# Combining all-metrics into one data-frame #
#############################################
FinalData <- base::rbind(MSE.Data, MAE.Data, PI.Data, WIS.Data)

###################
# Saving the data #
###################
utils::write.csv(FinalData, file = paste0("~/Georgia State R/MonkeyPox/Mpox Metrics Paper/Forecast-Metrics/skillScores.csv"), row.names = F)

```

`



