---
title: "Winkler Scores"
author: "Amanda Bleichrodt"
date: "2024-01-29"
output: html_document
---

# Needed packages 
```{r needed-packages}
library(tidyverse)
```

# Creating the function to calculate Winkler Scores
```{r winkler-function}
#################################
# Creating the Winkler Function #
#################################
winkler_score <- function(upper.bound, lower.bound, data){
  
  ##################################################
  # Runs if the observed data is lower than the LB #
  ##################################################
  if(data < lower.bound){
    
    # Calculated Winkler Score
    score <- (upper.bound - lower.bound) + (2/0.05)*(lower.bound-data)
  
  #################################################
  # Runs if the observed data falls in the bounds #
  #################################################
  }else if(data >= lower.bound & data <= upper.bound){
    
    # Calculated Winkler score
    score <- upper.bound - lower.bound
  
  ################################################
  # Runs if the observed data falls above the UB #
  ################################################
  }else{
    
    # Calculated Winkler score 
    score <- (upper.bound - lower.bound) + (2/0.05)*(data - upper.bound)
    
  }
  
  #######################
  # Returning the score #
  #######################
  return(score)
  
}
```

# Calculating Winkler Scores for ARIMA
```{r ARIMA-Forecasts-Winkler-Scores-1}
#############################################
# Empty data frame for ARIMA Winkler Scores #
#############################################
winker.scores.ARIMA <- data.frame(framework = NA, model = "ARIMA", weight = NA, area = NA, date = NA, horizon = NA, WS = NA)

#################################
# Creating a list of file-names #
#################################
ARIMA.forecasts <- list.files(
  path = paste0("Forecasts/ARIMA/Forecasts/Formatted"), 
          recursive = TRUE, # Indicates a pattern in files 
          pattern = paste0(".csv"), 
          full.names = T) # Calling full file names
```

## Applying the function to the ARIMA forecasts 
```{r ARIMA-Forecasts-Winkler-Scores-2}
###############################
# Looping through ARIMA files #
###############################
for(a in 1:length(ARIMA.forecasts)){
  
  # File Name 
  fileName <- ARIMA.forecasts[a]
  
  ######################################
  # Pulling information from file name #
  ######################################
  
  # Horizon
  horizon <- as.numeric(qdapRegex::ex_between(fileName, "horizon-", "-")[[1]])
  
  # Determining the location associated with the file
  area <- qdapRegex::ex_between(fileName, paste0("horizon-", horizon, "-"), "-")[[1]]
  
  # Determining the forecasting period associated with the file
  date <- anytime::anydate(qdapRegex::ex_between(fileName, paste0(area, "-"), "-Formatted")[[1]])
  
  # Calibration period
  calibration <- (qdapRegex::ex_between(fileName, "calibration-", "-horizon")[[1]])
  
  # Keeping only 11-weeks
  if(calibration != "11"){
    
    next
    
  }
  
  #######################################
  # Preparing the data for calculations #
  #######################################
  
  # Reading in the data
  data <- readr::read_csv(fileName)
  
  # Selecting the upper, lower, and observed data
  data.filters <- data %>%
    dplyr::mutate(Date = anytime::anydate(paste0(year, "-", month, "-", day))) %>% # Date column 
    dplyr::filter(Date > date) %>% # Keeping only forecasting horizon 
    dplyr::select(data, LB, UB) # Selecting needed variables
  
  ##################################
  # Calculating the Winkler Scores #
  ##################################
  
  # Creating vector for Winkler Scores
  scoreList <- NA
  
  # Looping through rows 
  for(h in 1:horizon){
          
  # Calculating the winkler scor
  scoreList[h] <- winkler_score(data.filters$UB[h], data.filters$LB[h], data.filters$data[h])
  
  }
  
  # Avg. Winkler Scores - Across forecasts 
  scoreTest <- mean(scoreList)
  
  ########################################
  # Preparing the data for later merging #
  ########################################
  winker.scores.ARIMA <- rbind(winker.scores.ARIMA, c(framework = NA, model = "ARIMA", weight = NA, area, as.character(date), horizon, scoreTest))
}
```

# Calculating Winkler Scores for the *n*-sub-epidemic model 
```{r n-sub-epidemic-Forecasts-Winkler-Scores-1}
############################
# Winkler Score Data-Frame #
############################
winker.scores.nse <- data.frame(framework = NA, model = NA, weight = NA, area = NA, date = NA, horizon = NA, WS = NA)

#################################
# Creating a list of file-names #
#################################
n.forecasts <- list.files(
  path = paste0("Forecasts/n-sub-epidemic/Calibration-11-Weeks"), 
          recursive = TRUE, # Indicates a pattern in files 
          pattern = paste0(".csv"), 
          full.names = T) # Calling full file names
```

## Applying the function to the *n*-sub-epidemic forecasts 
```{r n-sub-epidemic-Forecasts-Winkler-Scores-2}
########################################
# Looping through n-sub-epidemic files #
########################################
for(n in 1:length(n.forecasts)){
  
  # File Name 
  fileName <- n.forecasts[n]
  
  ######################################
  # Pulling information from file name #
  ######################################
  
  # Model 
  model <- qdapRegex::ex_between(fileName, "Weeks/", "-onsetfixed")[[1]][1]
  
  # Forecasting horizon
  horizon <- as.numeric(qdapRegex::ex_between(fileName, "horizon-", "-")[[1]][1])
  
  # Forecast date
  date <- str_sub(fileName,-14,-5)
  
  # Determining the location associated with the file
  area <- qdapRegex::ex_between(fileName, paste0("horizon-", horizon, "-"), paste0("-", date))[[1]]
          
  # Weight 
  if(grepl("Unweighted", model)){
    
   weight <- "-1"
    
  }else if(grepl("Weighted", model)){
    
    weight <- "1"
    
  }else{
    
    weight <- NA
    
  }
  
  #######################################
  # Preparing the data for calculations #
  #######################################
  
  # Reading in the data
  data <- readr::read_csv(fileName)
  
  # Selecting the upper, lower, and observed data
  data.filters <- data %>%
    dplyr::mutate(Date = anytime::anydate(paste0(year, "-", month, "-", day))) %>% # Date column 
    dplyr::filter(Date > anytime::anydate(date)) %>% # Keeping only forecasting horizon 
    dplyr::select(data, LB, UB) # Selecting needed variables
  
  ##################################
  # Calculating the Winkler Scores #
  ##################################
  
  # Creating vector for Winkler Scores
  scoreList <- NA
  
  # Looping through rows 
  for(h in 1:horizon){
          
  # Calculating the winkler scor
  scoreList[h] <- winkler_score(data.filters$UB[h], data.filters$LB[h], data.filters$data[h])
  
  }
  
  # Avg. Winkler Scores
  scoreTest <- mean(scoreList)
  
  ########################################
  # Preparing the data for later merging #
  ########################################
  winker.scores.nse <- rbind(winker.scores.nse, c(framework = "SE", model, weight, area, date = as.character(date), horizon, scoreTest)) 

}
```

# Calculating Winkler Scores for the spatial-wave model 
```{r spatial-wave-Forecasts-Winkler-Scores-1}
############################
# Winkler Score Data-Frame #
############################
winker.scores.sw <- data.frame(framework = "SW", model = NA, weight = NA, area = NA, date = NA, horizon = NA, WS = NA)

#################################
# Creating a list of file-names #
#################################
sw.forecasts <- list.files(
  path = paste0("Forecasts/Spatial-wave/Calibration-11-Weeks"), 
          recursive = TRUE, # Indicates a pattern in files 
          pattern = paste0(".csv"), 
          full.names = T) # Calling full file names
```

## Applying the function to the spatial-wave forecasts 
```{r spatial-wave-Forecasts-Winkler-Scores-2}
######################################
# Looping through spatial-wave files #
######################################
for(w in 1:length(sw.forecasts)){
  
  # File Name 
  fileName <- sw.forecasts[w]
  
  ######################################
  # Pulling information from file name #
  ######################################
  
  # Model 
  model <- qdapRegex::ex_between(fileName, "Weeks/", "-onsetfixed")[[1]][1]
  
  # Forecasting horizon
  horizon <- as.numeric(qdapRegex::ex_between(fileName, "horizon-", "-")[[1]][1])
  
  # Forecast date
  date <- str_sub(fileName,-14,-5)
  
  # Determining the location associated with the file
  area <- qdapRegex::ex_between(fileName, paste0("horizon-", horizon, "-"), paste0("-", date))[[1]]
          
  # Weight 
  if(grepl("Unweighted", model)){
    
   weight <- "-1"
    
  }else if(grepl("Weighted", model)){
    
    weight <- "1"
    
  }else{
    
    weight <- NA
    
  }
  
  #######################################
  # Preparing the data for calculations #
  #######################################
  
  # Reading in the data
  data <- readr::read_csv(fileName)
  
  # Selecting the upper, lower, and observed data
  data.filters <- data %>%
    dplyr::mutate(Date = anytime::anydate(paste0(year, "-", month, "-", day))) %>% # Date column 
    dplyr::filter(Date > anytime::anydate(date)) %>% # Keeping only forecasting horizon 
    dplyr::select(data, LB, UB) # Selecting needed variables
  
  ##################################
  # Calculating the Winkler Scores #
  ##################################
  
  # Creating vector for Winkler Scores
  scoreList <- NA
  
  # Looping through rows 
  for(h in 1:horizon){
          
  # Calculating the winkler scor
  scoreList[h] <- winkler_score(data.filters$UB[h], data.filters$LB[h], data.filters$data[h])
  
  }
  
  # Avg. Winkler Scores
  scoreTest <- mean(scoreList)
  
  ########################################
  # Preparing the data for later merging #
  ########################################
  winker.scores.sw <- rbind(winker.scores.sw, c(framework = "SW", model, weight, area, date = as.character(date), horizon, scoreTest)) 

}
```

# Combining the Winkler Scores - Crude
```{r crude-skill-scores}
###############################
# Combining the crude metrics #
###############################
all.avg.winkler <- base::rbind(winker.scores.ARIMA[-1,] , winker.scores.nse[-1,], winker.scores.sw[-1,]) %>%
  dplyr::mutate(Model = ifelse(!is.na(framework), paste0(framework, "-", model), model), # Model variable 
                Location = area) %>% # Location
  dplyr::ungroup() %>% # Ungrouping 
  dplyr::select(Location, Model, date, horizon, WS) # Selecting needed variables

###########################################
# Saving the date-specific Winkler scores #
###########################################
write.csv(all.avg.winkler, "Forecast-Metrics/All-winkler-Scores.csv", row.names = F)

```

# Combining the Winkler Scores - Average 
```{r combining-winkler-scores}
#################################
# Finding the avg. WS for ARIMA #
#################################
average.ARIMA <- winker.scores.ARIMA[-1,] %>%
  dplyr::group_by(framework, model, weight, area, horizon) %>% # Grouping by variables other than date and scoreTest
  dplyr::mutate(WS = as.numeric(WS), # Changing winkler score to numeric 
                meanWS = mean(WS),
                date = anytime::anydate(date)) # Mean of winkler scores across forecasting periods 

###############################
# Finding the avg. WS for NSE #
###############################
average.NSE <- winker.scores.nse[-1,] %>%
  dplyr::group_by(framework, model, weight, area, horizon)  %>% # Grouping by variables other than date and scoreTest
  dplyr::mutate(WS = as.numeric(WS), # Changing winkler score to numeric 
                meanWS = mean(WS),
                date = anytime::anydate(date)) # Mean of winkler scores across forecasting periods 

##############################
# Finding the avg. WS for SW #
##############################
average.SW <- winker.scores.sw[-1,] %>%
  dplyr::group_by(framework, model, weight, area, horizon)  %>% # Grouping by variables other than date and scoreTest
  dplyr::mutate(WS = as.numeric(WS), # Changing winkler score to numeric 
                meanWS = mean(WS),
                date = anytime::anydate(date)) # Mean of winkler scores across forecasting periods 

#############################
# Combining average metrics #
#############################
all.avg.winkler <- base::rbind(average.ARIMA, average.NSE, average.SW)
```

## Cleaning the data set 
```{r cleaning-up-average-metrics}
#################################
# Cleaning of the avg. data set #
#################################
cleanAvg <- all.avg.winkler %>%
  ungroup() %>% # Ungrouping 
  dplyr::distinct(framework, model, area, horizon, .keep_all = T) %>% # Needed variables
  
  ######################################
  # Creating the combined model column #
  ######################################
  dplyr::mutate(Location = area, # Location 
                model_final = ifelse(is.na(framework) & is.na(weight), 
                                     model,
                                     paste0(framework, "-", model))) %>%
                                  
  ##################################   
  # Selecting the needed variables #
  ##################################
  dplyr::select(Location, model, horizon, model_final, meanWS)

##############################
# Saving the average metrics #
##############################
write.csv(cleanAvg, "Forecast-Metrics/Average-winkler-Scores.csv", row.names = F)
```

